---
title: "Overview of S-map coefficient results"
author: "Koya Hashimoto"
date: "2023/10/6"
output:
  md_document:
    variant: markdown_github
---

# Overview of S-map coefficient results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load required packages

```{r library}
sessionInfo() #save session information (R version 3.6.3 (2020))

library(ggplot2) #3.3.2
library(dplyr) #1.0.2
library(tidyr) #1.1.2
library(patchwork) #1.1.1
library(igraph) #1.2.6
library(RColorBrewer) #1.1.2
```

### Data compilation

```{r data}
all_multsmap_coefs <- read.csv("./processed_data/all_regulsmap_coefs240123.csv")
head(all_multsmap_coefs)

all_multsmap_coefs$Recipient <- 
  factor(all_multsmap_coefs$Recipient, 
         levels=unique(all_multsmap_coefs$Recipient))
all_multsmap_coefs$Donor <- 
  factor(all_multsmap_coefs$Donor, 
         levels=c("Phytopla1", "Phytopla1_1", "Roti1", "Roti1_1", "Roti1_2", 
                  "Zoopla1", "Zoopla1_1", "Zoopla1_2", "Zoopla1_3", "Zoopla1_4", 
                  "Mp1", "Mp1_1", "Det1", "Herb1", "Herb1_1", "Herb1_2", 
                  "Pred.P1", "Pred.P1_1", "Pred.P1_2", "Pred.P1_3", "Pred.B1", 
                  "Pred.S1", "Pred.S1_1", "Pred.S1_2", "Pred.S1_3", "Moll1", "0"))
all_multsmap_coefs$Treatment <- factor(all_multsmap_coefs$Treatment, levels=c("Cont", "Fipro", "Pent", "Joint"))
```

## Overview

### Furequency distribution of mean S-map coefs

```{r IS distribution}
all_multsmap_coefs <- 
  all_multsmap_coefs %>%
  unite(col="Recipient_Donor", Recipient, Donor, remove=FALSE)
all_multsmap_coefs_smrz <- 
  all_multsmap_coefs %>%
  group_by(Treatment, Recipient, Donor, r_d, Recipient_Donor, Year, Tank) %>%
  summarise(smap_coef=mean(smap_coef, na.rm=TRUE)) %>%
  group_by(Treatment, Recipient, Donor, r_d, Recipient_Donor) %>%
  summarise(smap_coef=mean(smap_coef, na.rm=TRUE))

all_multsmap_coefs_smrz_intersp <- 
  all_multsmap_coefs_smrz %>%
  separate(Donor, into=c("Donor_", "delay"), sep="_", remove=FALSE) %>%
  filter(Recipient!=Donor_ & r_d!="c_0") %>%
  select(-Donor_, - delay)

all_multsmap_coefs_smrz_intersp_abs <- 
  all_multsmap_coefs_smrz_intersp %>%
  mutate(Sign=ifelse(smap_coef>0, "Positive", "Negative"), abs_smap_coef=abs(smap_coef)) %>%
  arrange(Treatment, Recipient, Donor)

f_labels <- data.frame(Treatment=c("Cont", "Fipro", "Pent", "Joint"), 
                       label=rep("↓", times=4), 
                       x=tapply(all_multsmap_coefs_smrz_intersp_abs$abs_smap_coef, 
                                all_multsmap_coefs_smrz_intersp_abs$Treatment, 
                                function(x) median(x, na.rm=TRUE)), 
                       y=rep(7, times=4))
ggplot(all_multsmap_coefs_smrz_intersp_abs) +
  geom_histogram(position="stack", binwidth=0.05, mapping=aes(x=abs_smap_coef, fill=Sign)) + facet_wrap(~Treatment) +
  geom_text(data=f_labels, aes(label=label, x=x, y=y), size=10)
```

### Networks

*Orange: negative effect. Blue: positive effect. Thickness: mean interaction strength. Dashed arrow: 0.05<P<0.1.

```{r interaction network, fig.width=9, fig.height=8}
library(igraph)
g1 <- graph(edges=c(2,1, 5,1, 7,1,
                    1,2, 6,2, 7,2,
                    2,3,
                    2,4, 6,4, 7,4,
                    1,5, 2,5, 4,5,
                    1,6, 4,6, 5,6,
                    2,7,
                    4,9,
                    2,10, 3,10, 6,10, 8,10, 9,10), n=10, directed=T)
IS.Tr.mean <- all_multsmap_coefs_smrz_intersp_abs
V(g1)$name <- c("Phytopla", "Rotifer", "Zoopla.Cru", "Macrophyte", "Detritivore", "Herbivore",
                "Phyto.predator", "Bent.predator", "Surf.predator", "Mollusca") #save nodes' names
l <- layout_in_circle(g1) #layout in circle

#each treatment
par(mfrow=c(2, 2), mar=c(3, 3, 3, 3), family="sans")
for (i in c("Cont", "Fipro", "Pent", "Joint")){
  plot(g1, layout=l, vertex.label.family="sans",
       edge.lty=c(1, 2, 1,
                  1, 1, 1,
                  1,
                  1, 1, 1,
                  1, 2, 1,
                  1, 1, 1,
                  1,
                  1,
                  1, 1, 1, 1, 1),
       edge.color=c(ifelse(subset(IS.Tr.mean, Treatment==i)[,"Sign"]=="Positive", 2, 6)),
       edge.curved=0.1,
       edge.width=subset(IS.Tr.mean, Treatment==i)$abs_smap_coef*20)
  legend("topleft", legend=i, bty="n", x.intersp=0, y.intersp=0, cex=1.2)
  if (i=="Fipro") text(c(-0.98, -0.44), c(-0.73, -1.14), labels="*", cex=4) else {
    if (i=="Pent") text(-0.42, 1.12, labels="*", cex=4) else { 
      if (i=="Joint") text(c(-0.98, -0.44, 0.92, 1.1, -0.42, -1.09),
                           c(-0.73, -1.14, -0.73, -0.11, 1.12, -0.11),
                           labels=c("*", "*", "**", "*", "*", "*"), cex=4)
    }
  }
  
}
```

## Generate Fig. S1

```{r suppl, fig.width=7, fig.height=10}

palSet1 <- brewer.pal(9, "Set1")

g1 <- graph(edges=c(2,1, 5,1, 7,1, 
                    1,2, 6,2, 7,2, 
                    2,3, 
                    2,4, 6,4, 7,4, 
                    1,5, 2,5, 4,5,
                    1,6, 4,6, 5,6, 
                    2,7, 
                    4,9,
                    2,10, 3,10, 6,10, 8,10, 9,10), n=10, directed=T)

V(g1)$name <- c("Phytoplankton", "Rotifers", "Crustacean\nzooplankton", "Macrophytes", 
                "Detritivores", "Herbivores", 
                "Phytophilous\npredators", "Benthic\npredators", "Neustonic\npredators", "Molluscs") #save nodes' names
l <- layout_in_circle(g1) #layout in circle

#windows(7, 10, rescale="fixed")
#pdf("./figs/figS1.pdf", width=7, height=10)
layout(matrix(c(1, 2, 5, 6, 3, 4, 7, 8), ncol=2, byrow=TRUE), heights=c(2.5, 1, 2.5, 1))
#each treatment
par(mar=c(2, 2, 2, 2), family="sans")
plot(g1, layout=l, vertex.label.family="sans",
     edge.lty=c(1, 2, 1, 
                1, 1, 1,
                1, 
                1, 1, 1, 
                1, 2, 1, 
                1, 1, 1, 
                1, 
                1, 
                1, 1, 1, 1, 1), 
     edge.color=c(ifelse(subset(IS.Tr.mean, Treatment=="Cont")[,"Sign"]=="Positive", palSet1[2], palSet1[1])), 
     edge.curved=0.1, 
     edge.width=subset(IS.Tr.mean, Treatment=="Cont")$abs_smap_coef*20, 
     vertex.label.dist=1.5, vertex.label.degree=-pi/2, 
     vertex.label.color="black")
legend("topleft", legend="a", bty="n", xpd=NA, x.intersp=-1, y.intersp=-1, cex=1.5, text.font=2)
legend("topleft", legend="C", bty="n", x.intersp=1, y.intersp=1, cex=1.5, text.font=2)

plot(g1, layout=l, vertex.label.family="sans",
     edge.lty=c(1, 2, 1, 
                1, 1, 1,
                1, 
                1, 1, 1, 
                1, 2, 1, 
                1, 1, 1, 
                1, 
                1, 
                1, 1, 1, 1, 1), 
     edge.color=c(ifelse(subset(IS.Tr.mean, Treatment=="Fipro")[,"Sign"]=="Positive", palSet1[2], palSet1[1])), 
     edge.curved=0.1, 
     edge.width=subset(IS.Tr.mean, Treatment=="Fipro")$abs_smap_coef*20, 
     vertex.label.dist=1.5, vertex.label.degree=-pi/2, 
     vertex.label.color="black")
legend("topleft", legend="b", bty="n", xpd=NA, x.intersp=-1, y.intersp=-1, cex=1.5, text.font=2)
legend("topleft", legend="I", bty="n", x.intersp=1, y.intersp=1, cex=1.5, text.font=2)
text(c(-0.98, -0.44), c(-0.73, -1.14), labels="*", cex=4)

plot(g1, layout=l, vertex.label.family="sans",
     edge.lty=c(1, 2, 1, 
                1, 1, 1,
                1, 
                1, 1, 1, 
                1, 2, 1, 
                1, 1, 1, 
                1, 
                1, 
                1, 1, 1, 1, 1), 
     edge.color=c(ifelse(subset(IS.Tr.mean, Treatment=="Pent")[,"Sign"]=="Positive", palSet1[2], palSet1[1])), 
     edge.curved=0.1, 
     edge.width=subset(IS.Tr.mean, Treatment=="Pent")$abs_smap_coef*20, 
     vertex.label.dist=1.5, vertex.label.degree=-pi/2, 
     vertex.label.color="black")
legend("topleft", legend="e", bty="n", xpd=NA, x.intersp=-1, y.intersp=-1, cex=1.5, text.font=2)
legend("topleft", legend="H", bty="n", x.intersp=1, y.intersp=1, cex=1.5, text.font=2)
text(-0.42, 1, labels="*", cex=4)

plot(g1, layout=l, vertex.label.family="sans",
     edge.lty=c(1, 2, 1, 
                1, 1, 1,
                1, 
                1, 1, 1, 
                1, 2, 1, 
                1, 1, 1, 
                1, 
                1, 
                1, 1, 1, 1, 1), 
     edge.color=c(ifelse(subset(IS.Tr.mean, Treatment=="Joint")[,"Sign"]=="Positive", palSet1[2], palSet1[1])), 
     edge.curved=0.1, 
     edge.width=subset(IS.Tr.mean, Treatment=="Joint")$abs_smap_coef*20, 
     vertex.label.dist=1.5, vertex.label.degree=-pi/2, 
     vertex.label.color="black")
legend("topleft", legend="f", bty="n", xpd=NA, x.intersp=-1, y.intersp=-1, cex=1.5, text.font=2)
legend("topleft", legend="I+H", bty="n", x.intersp=1, y.intersp=1, cex=1.5, text.font=2)
text(c(-0.98, -0.44, 0.92, 1.1, -0.42, -1.09), 
     c(-0.73, -1.14, -0.73, -0.11, 1, -0.11), 
     labels=c("*", "*", "**", "*", "*", "*"), cex=4)


par(mar=c(4, 2, 0, 2), mgp=c(2.5, 1, 0), las=1, family="sans", cex=0.8)
for (i in 1:4) {
  hist(subset(IS.Tr.mean, Treatment==c("Cont", "Fipro", "Pent", "Joint")[i])$abs_smap_coef, breaks=seq(0, 0.5, 0.05),
       xlab="Mean interaction strength", col=palSet1[1], main=c("")[i])
  lines(hist(subset(IS.Tr.mean, Treatment==c("Cont", "Fipro", "Pent", "Joint")[i] & Sign=="Positive")$abs_smap_coef, plot=FALSE, 
             breaks=seq(0, 0.5, 0.05)), col=palSet1[2], xlim=c(0, 0.5))
  arrows(median(subset(IS.Tr.mean, Treatment==c("Cont", "Fipro", "Pent", "Joint")[i])$abs_smap_coef), 9, 
         median(subset(IS.Tr.mean, Treatment==c("Cont", "Fipro", "Pent", "Joint")[i])$abs_smap_coef), 8, code=2, length=0.05, lwd=2, xpd=NA)
  legend("topright", legend=c("Negative interactions", "Positive interactions"), bty="n",
         pt.bg=palSet1[1:2], pch=22, plot=c(TRUE, FALSE, FALSE, FALSE)[i], pt.cex=2)
  legend("topleft", legend=c("c", "d", "g", "h")[i], 
         bty="n", xpd=NA, x.intersp=-1.2, y.intersp=-2, cex=1.2, text.font=2)
}
dev.off()
```

## Source of variation in the S-map coefs (Fig. S4) 

We performed two-way ANOVA for every interaction link, examining the effects of treatment, census date and their interaction on S-map coefficient values, to confirm whether the temporal fluctuations of the interaction effect were consistent among replicates. The mean squares of residuals are unbiased estimates of the variance of various noises including differences in individual replicates. If these values are small relative to those of census dates or treatments, temporal fluctuations of s-map coefs should be considered as relatively consistent among replicates.

```{r smap coef source of variation, fig.width=6, fig.height=6}
all_multsmap_coefs <- 
  all_multsmap_coefs %>%
  unite(col="Recipient_Donor", Recipient, Donor, remove=FALSE)

anova_res <- 
  all_multsmap_coefs %>%
  unite("Year_Week", Year, Week, sep="_", remove=FALSE) %>%
  split(., .$Recipient_Donor) %>%
  lapply(., function(data) {
    res <- as.data.frame(anova(lm(smap_coef ~ Treatment * Year_Week, data)))
    res <- cbind(Source=rownames(res), res)
    res}) %>%
  do.call(rbind, .) %>% 
  cbind(., Recipient_Donor=rep(unique(all_multsmap_coefs$Recipient_Donor), each=4))

theme_set(theme_bw())
anova_res$Source <- factor(anova_res$Source, levels=unique(anova_res$Source))
ggplot(anova_res, aes(y=!!as.name("Mean Sq"), x=Source, group=Recipient_Donor)) + 
  geom_point() + geom_line() + scale_y_log10() + 
  ylab("Mean squares") + scale_x_discrete(labels=c("Treatment"="Treatment", 
                                                   "Year_Week"="Census date", 
                                                   "Treatment:Year_Week"="Treatment x Census date", 
                                                   "Residuals"="Residuals"))
ggsave("./figs/figS4.ggplot2.pdf", width=6, height=6, device=cairo_pdf, unit="in")

ggplot(anova_res, aes(y=!!as.name("F value"), x=Source, group=Recipient_Donor)) + 
  geom_point() + geom_line() + scale_y_log10()

F_value_count <- matrix(c(nrow(subset(anova_res, Source=="Treatment" & anova_res[,"F value"]<=1)),
                          nrow(subset(anova_res, Source=="Treatment" & anova_res[,"F value"]>1)),
                          nrow(subset(anova_res, Source=="Year_Week" & anova_res[,"F value"]<=1)),
                          nrow(subset(anova_res, Source=="Year_Week" & anova_res[,"F value"]>1)),
                          nrow(subset(anova_res, Source=="Treatment:Year_Week" & anova_res[,"F value"]<=1)),
                          nrow(subset(anova_res, Source=="Treatment:Year_Week" & anova_res[,"F value"]>1))), nrow=3, byrow=TRUE)
rownames(F_value_count) <- c("Treatment", "Year_Week", "Treatment:Year_Week")
colnames(F_value_count) <- c("F<=1", "F>1")
F_value_count
```
